---
title: "ghgemissions"
---

## Clean State GHG data

The [EPA's State GHG Emissions and Removals dataset](https://www.epa.gov/ghgemissions/state-ghg-emissions-and-removals) contains state-level estimates of greenhouse gas emissions between 1990-2020.

The data is in wide format, meaning the emissions for each year are contained in a column e.g. `Y2018`, `Y2019`, `Y2020`, etc.
To analyze the data, it is useful to pivot the data so that each yearly estimate is in its own row rather than a column. 

```{r}
library(tidyverse)
library(readxl)
library(here)
library(janitor)

state_file = here("data","AllStateGHGData_042623.xlsx")

state_ghg = read_excel(state_file, sheet = "Data by UNFCCC-IPCC Sectors", guess_max = 18000) |> 
  janitor::clean_names()
glimpse(state_ghg)
```

Pivot the year columns to `year` and store the values in `emissions`.
I use the `names_transform` argument to remove the 'y' prefix and convert the result to an integer.
Also add state and territory names.

```{r}
state_ids = tibble(state = state.abb, state_name = state.name)

(state_ghg_pivot = state_ghg |> 
   left_join(state_ids, by = "state") |> 
    mutate(state_name = case_when(state == "AS" ~ "American Samoa",
                                  state == "DC" ~ "Washington DC",
                                  state == "FM" ~ "Federated States of Micronesia",
                                  state == "FO" ~ "Federal Offshore",
                                  state == "GU" ~ "Guam",
                                  state == "MH" ~ "Marshall Islands",
                                  state == "MP" ~ "Northern Mariana Islands",
                                  state == "PR" ~ "Puerto Rico",
                                  state == "PW" ~ "Palau",
                                  state == "UM" ~ "U.S. Minor Outlying Islands",
                                  state == "VI" ~ "U.S. Virgin Islands",
                                  is.na(state_name) ~ state,
                                  T ~ state_name)) |> 
  pivot_longer(cols = matches("y[0-9]{4}"), names_to = "year", 
               names_transform = list(year = \(x) as.integer(str_remove_all(x, "^y"))),
               values_to = "emissions")) |> 
  glimpse()
```
Write the pivoted data to a csv file.

```{r}
write_csv(state_ghg_pivot, file = here("data", "AllStatePivot.csv"), na = "")

```

